---
title: "Week 1 assignment"
output: html_document
date: '2022-03-20'
---

# Author: Chander Lekha

#### Exercises

#1.  *Explain coercion rules. Create an example that demonstrates the fixed order of the coercion rules.*

# The coercion rules are [logical → integer → numeric → character]
```{r}
rand_logicalVal <- runif(10) <0.7

typeof(rand_logicalVal)
is.logical(rand_logicalVal)

int_val <- as.integer(rand_logicalVal)

typeof(int_val)
is.integer(int_val)

numeric_val <- as.numeric(int_val)

typeof(numeric_val)
is.numeric(numeric_val)

char_val <- as.character(numeric_val)

typeof(char_val)
is.character(char_val)

```

#2.  *The following code will generate a warning. Explain why.* 

```{r}
as.integer(c("one", "two", "three"))

#### This does not follow the rule of coercion.This will not force non numeric/integer strings to integers.  The coercion rules are [logical → integer → numeric → character]
```

#3.  *Create a vector of a sequence of integers starting at 2, ending at 100, and increasing by 2. Assign the vector to an object called `evens`. Coerce the vector to be of type `integer`. Confirm that the vector is of type `integer`.*

```{r}
evens <- seq(from = 2L, to = 100L , by = 2L )

typeof(evens)

```

#4.  *Create a vector of random numbers with a length equal to your vector in exercise 3. Use a seed with a value of `730` before creating your vector. Assign the vector to an object called `randoms`. Confirm the type of the vector.*

```{r}

set.seed(730)
randoms <- seq(  from = 2, to = 100  , by = 2 )

typeof(randoms)
```

#5.  *Create a vector demonstrating which values in exercise 4 are greater than `0.75`. Assign the vector to an object called `logicals`. Confirm the type of the vector.*

```{r}
logicals <- c(2:100,2) < 0.75

typeof(logicals)
print(logicals)
```

#6.  Create a data frame including the three vectors created in exercises 3 through 5. Assign it to an object following best practices for naming.

```{r}
evens <- seq(from = 2L, to = 100L , by = 2L )
randoms <- seq(  from = 2, to = 100  , by = 2 )
logicals <- c(2:100,2) < 0.75

vec.data <- data.frame(even_num = evens,
           randoms_data = randoms,
           logical_data = logicals
)
```

#7.  *Create a new object which is the sum of `evens` and `randoms` all divided by 3. Demonstrate how to calculate basic summary statistics for this new object (mean, median, minimum, maximum, standard deviation, variance).*

```{r}
evens <- seq(from = 2L, to = 100L , by = 2L )
randoms <- seq(  from = 2, to = 100  , by = 2 )

totalOfEvensAndRandoms <- (evens + randoms)/3

print(totalOfEvensAndRandoms)

mean(totalOfEvensAndRandoms)

median(totalOfEvensAndRandoms); min(totalOfEvensAndRandoms); max(totalOfEvensAndRandoms); var(totalOfEvensAndRandoms)
sd(totalOfEvensAndRandoms)
```

# 8.  *We discussed the idea that R is a "vectorized" language. Explain this concept. Why is it important?*
### R is used to perform the very same operation on different data sets simultanously.



# 9.  *Run the `sessionInfo` function in a chunk.*

```{r}
sessionInfo(package = NULL)
```





