---
title: "Class_01_Content"
author: "Ryan Harrington"
date: "3/14/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Using R as a calculator

This is some text about calculators!

```{r}
2 + 4
```

```{r}
# I am a comment
# This text will not be executed
1 - 4

7 * 4

9 / 3

8 + 4 - 3

6 / (1 + 2)         # here's a comment as well
```

# Making comparisons

```{r}
3 < 5

# True is evaluated as...
TRUE 
T

## but not this
# true

# False is evaluated as...
FALSE
F

## but not this
# false

```

```{r}
# < <= > >=

4 >= 6
3.5 == 2.1

# ==
1 == 2

7 ^ 2 >= 2 ^ 6 - 4 ^ 2
```

# Object assignment

```{r}
value <- 6 + 3
```

```{r}
value + 3

value ^ 2

value / 3 == sqrt(value)
```

```{r}
## This won't work because it doesn't match the case

# Value
```

```{r}
interesting_value <- 2 ^ 6 - 4 ^ 2
```

```{r}
# Use arrows
another_value <- 6 + 2

# Not ='s
a_third_value = 7 - 2

# But they both work
```

```{r}
another_value + a_third_value
```

```{r}
.variable_3 <- another_value
```

# Creating vectors

```{r}
1:10

21:72

10:1

-3:-8
```

```{r}
21:72 + 3
```

```{r}
top_ten <- 1:10
```

```{r}
top_ten / 8
```

```{r}
top_ten == 8
```

```{r}
1 : 10
```

# Functions

```{r}
sqrt(3)

abs(-2)
```

```{r}
sqrt(top_ten)
```

```{r}
mean(top_ten)
median(top_ten)
min(top_ten)
max(top_ten)
sd(top_ten)
var(top_ten)
```

# Making vectors

```{r}
lost_numbers <- c(4, 8, 15, 16, 23, 42)
```

```{r}
jennys_number <- c(8, 6, 7, 5, 3, 0, 9)
```

```{r}
transcendental_numbers <- c(pi, 1.618, 2 * pi, 2.718)
```

```{r}
new_sequence <- c(1:10, 8, 6, 7, 5, 3, 0, 9)
```

```{r}
all_sequences <- c(lost_numbers, transcendental_numbers, top_ten)
```

```{r}
seq(1, 10)
```

```{r}
seq(1, 10, 1)
```

```{r}
seq(1, 10, 2)
```

```{r}
seq(from = 1, to = 10, by = 2)
```

```{r}
odd_numbers_10 <- seq(to = 10, from = 1, by = 2)
```

```{r}
dice <- 1:6

sample(dice, size = 20, replace = TRUE)
```

```{r}
coin <- c("H", "T")

sample(coin, size = 10, replace = TRUE)
```

# Data Types

## Doubles and Integers

```{r}
lost_numbers
```

```{r}
transcendental_numbers
```

```{r}
typeof(lost_numbers)

is.double(lost_numbers)
is.integer(lost_numbers)
is.numeric(lost_numbers)
```

```{r}
lost_numbers <- as.integer(lost_numbers)

typeof(lost_numbers)

is.double(lost_numbers)
is.integer(lost_numbers)
is.numeric(lost_numbers)
```

```{r}
is.integer(1)
is.integer(1L)

is.integer(42L)

typeof(1L:10L)


typeof(1:10L)
typeof(1L:10)
```


## Logical

```{r}
set.seed(112)

random_logicals <- runif(137) < 0.5
```

```{r}
typeof(random_logicals)

is.logical(random_logicals)
```

```{r}
as.numeric(random_logicals)
```

How many items in the `random_logicals` list are `TRUE`?

```{r}
sum(random_logicals)
```

What percentage of items in the `random_logicals` list are `TRUE`?

```{r}
mean(random_logicals)
```


## Character

```{r}
months_of_the_year <- c("January", "February", "March", "April", "May", "June",
                        "July", "August", "September", "October", "November", "December")
```

```{r}
typeof(months_of_the_year)

is.character(months_of_the_year)
```

```{r}
as.numeric(months_of_the_year)
```

```{r}
as.character(random_logicals)
```


## Coercion

logical → integer → numeric → character

```{r}
random_logicals
is.logical(random_logicals)

random_logicals_ints <- as.integer(random_logicals)
is.integer(random_logicals_ints)

random_logicals_dbls <- as.double(random_logicals_ints)
is.double(random_logicals_dbls)

random_logicals_char <- as.character(random_logicals_ints)
is.character(random_logicals_char)
```


```{r}
as.raw(40)
charToRaw("A test string")
```

# Data Types

## Vectors

Everything we've done so far is a vector

## Lists

```{r}
random_logicals

months_of_the_year
```

```{r}
combined_list <- list(random_logicals, months_of_the_year)
```

```{r}
combined_list[[1]]
```

```{r}
combined_list[[2]]
```

## Data Frames

```{r}
months_df <- 
  data.frame(
    months = months_of_the_year
    ,letters = LETTERS[1:12]
    # ,month_number = 1:12
  )


months_df <- 
  data.frame(
    months = months_of_the_year,
    letters =          LETTERS[1:12]#,
    # month_number = 1:12
  )
```

```{r}
data.frame(months = months_of_the_year,
             month_number = 1:11)
```

